---
title: "Pupil Trajectories in Response to Light Stimulus --  FDA Analysis"
author: "SG"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: true
    theme: lumen
---

# Background
Pupil size light reflex to a light test is a potential test to determine a person is under the influence of THC and be able to be use as a field sobriety measures. The goals of the analysis to: 

## Potential Topics
1. Determine the variability of pupil light reflex in a sober population
2. Determine the effectiveness of pupil light reflex to measure sobriety after smoking THC 
    Question - dose response?
3. Jointly model pupil light reflex and blink rate during a light test (another data set)

Pupil size reflex to light does not habituated to THC use (smoking).

Time for smoking to post test maybe an important variable to model b/c: 

+ effects of THC reduce with time
+ currently time from smoke period to post are all similar so we might not see an effect

There is a circadian pattern to pupil size (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7445830/). 

## Experimental Design
+ Setting: Office
  + Were the lights ON/OFF during testing (JW)

## Features of Importance
1. Point of minimal constriction -- more dilation after smoking
2. Rebound dilation
+ Ideally the non-user data should have little change from pre to post but there may be some "habituation" to the test?

## Data summary
Data on pupil size during light reflex test. Pupil size was extracted at the image level based on image analysis techniques. Each test was performed simultaneously on right and left eye before and after THC use (smoking).

+ USE percent change for analysis
+ Start at frame 0 
+ End of test is not strictly defined (with FoS we shouldn't need to define the end of the test)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openxlsx)
library(ggplot2)
library(table1)
library(refund)
library(refund.shiny)
library(pROC)
library(stringr)
library(PerformanceAnalytics)
library(mgcv)
library(grid)
library(gridExtra)

# ps_folder <- "C:/Repositories/Dissertation/PupilSize_img"
# data_folder <- "C:/Users/Suneeta/OneDrive - The University of Colorado Denver/FDA/Data" # Home laptop
data_folder <- "C:/Users/godbolsu/OneDrive - The University of Colorado Denver/FDA/Data" # Work laptop
# data_folder <- "C:/Users/godbolsu/OneDrive - The University of Colorado Denver/FDA/Data"
ps_folder <- "202208_PupilLightReflex"
```

# Data Preparation
```{r dataReadIn, echo=FALSE}
ps <- read.csv(file.path(data_folder, ps_folder, "pupil_data_with_demo_20220926.csv"))

ps$demo_gender <- factor(ps$demo_gender, 
                         levels = c(1,2), 
                         labels = c("Male", "Female"))

ps$user_cat <- NA
ps$user_cat[ps$user_type == "non-user"] <- 0
ps$user_cat[ps$user_type == "occasional"] <- 1
ps$user_cat[ps$user_type == "daily"] <- 2

ps$user_cat <- factor(ps$user_cat, 
                      levels = c(0,1,2), 
                      labels = c("non-user", 
                                 "occasional", 
                                 "daily"))

ps$tp <- NA
ps$tp[ps$time == "pre2"] <- 0
ps$tp[ps$time == "post"] <- 1

ps$tp <- factor(ps$tp, 
                levels = c(0,1), 
                labels = c("pre2", "post"))
```

```{r pt_summ, echo=FALSE}
# total number of subjects
# length(unique(ps$subject_id))

# subject level data 
pt.df <- unique(ps[, c("subject_id", "tp", "user_cat", "demo_age", 
                "demo_weight", "demo_height", "demo_gender", "thc")])

# # Demographics Table by User Category
# table1(~ demo_age + demo_weight + demo_height + demo_gender|user_cat, 
#        data = pt.df[pt.df$tp == "pre2", ])

# number of subjects by timepoint (pre/post)
# table(pt.df$tp)
```

```{r scalarFeat, echo=FALSE}
scalar.feat <- read.csv(file.path(data_folder, ps_folder, "scalars_trim.csv"), 
                        header = T, stringsAsFactors = F)

scalar.feat$subject_id <- substr(scalar.feat$timeid, 1, 7)
scalar.feat$tp <- substr(scalar.feat$timeid, 8, 11)

scalar.featR <- scalar.feat[scalar.feat$eye == "Right", ]

pt.df <- merge(pt.df, scalar.featR[, c("subject_id", "tp", "min_constriction", 
                                       "duration", "avg_end_percent", "slope", 
                                       "AUC", "eye")], 
               by = c("subject_id", "tp"))

```

```{r ptdf_reshape, echo=FALSE}
pt.df.w <- reshape(pt.df, 
                   timevar = "tp", 
                   idvar = c("subject_id", "user_cat", "demo_age", 
                             "demo_weight", "demo_height", "demo_gender", "eye"), 
                   direction = "wide")
pt.df.w$thc.post[pt.df.w$user_cat == "non-user" & is.na(pt.df.w$thc.post)] <- 0
pt.df.w$thc_chg <- pt.df.w$thc.post - pt.df.w$thc.pre2
pt.df.w$BMI <- pt.df.w$demo_weight*703/(pt.df.w$demo_height)^2
pt.df.w$min_constriction_chg <- pt.df.w$min_constriction.post - pt.df.w$min_constriction.pre2
pt.df.w$AUC_chg <- pt.df.w$AUC.post - pt.df.w$AUC.pre2
pt.df.w$duration_chg <- pt.df.w$duration.post - pt.df.w$duration.pre2
pt.df.w$avg_end_percent_chg <- pt.df.w$avg_end_percent.post - pt.df.w$avg_end_percent.pre2
pt.df.w$slope_chg <-  pt.df.w$slope.post - pt.df.w$slope.pre2
```


```{r readIn_testTimes, echo=FALSE}
## Need to work on; time formatted in Excel file.
testTimes <- read.xlsx(file.path(data_folder, ps_folder, "All SafetyScan Times_23Aug2022_revSG.xlsx"), 
                       sheet = "Raw")

testTimes$pre_safetyscan_date_convert <- as.Date(testTimes$pre_safetyscan_date, origin = "1899-12-30")

testTimes$pre_safetyscan_time_convert <- as.POSIXct(paste0(testTimes$pre_safetyscan_date_convert, " ", testTimes$pre_safetyscan_time_hr, ":", testTimes$pre_safetyscan_time_min, ":", "00"), format = "%Y-%m-%d %H:%M:%S")

testTimes$consump_start_time_convert <- as.POSIXct(paste0(testTimes$pre_safetyscan_date_convert, " ", testTimes$consump_start_time_hr, ":", testTimes$consump_start_time_min, ":", "00"), format = "%Y-%m-%d %H:%M:%S")

testTimes$consump_end_time_convert <- as.POSIXct(paste0(testTimes$pre_safetyscan_date_convert, " ", testTimes$consump_end_time_hr, ":", testTimes$consump_end_time_min, ":", "00"), format = "%Y-%m-%d %H:%M:%S")

testTimes$post_safetyscan_time_convert <- as.POSIXct(paste0(testTimes$pre_safetyscan_date_convert, " ", testTimes$post_safetyscan_time_hr, ":", testTimes$post_safetyscan_time_min, ":", "00"), format = "%Y-%m-%d %H:%M:%S")

testTimes$postConsumpTimeToTest <- as.numeric(difftime(testTimes$post_safetyscan_time_convert,
                                                       testTimes$consump_end_time_convert,
                                                       units = "mins"))
testTimes$remove <- ifelse(testTimes$subject_id == "001-056" & is.na(testTimes$postConsumpTimeToTest), 1, 0)
testTimes <- testTimes[testTimes$remove == 0, ]

pt.df <- merge(pt.df.w, testTimes[, c("subject_id", "postConsumpTimeToTest")], 
               by = "subject_id")
ps <- merge(ps, testTimes[, c("subject_id", "postConsumpTimeToTest")], 
               by = "subject_id")

non_user.id <- pt.df$subject_id[pt.df$user_cat == "non-user"]
# View(testTimes[testTimes$subject_id %in% non_user.id, ])
```

## Table 1
No Consumption end time recorded for non-users
```{r tab1, echo=FALSE}
# Demographics Table by User Category
table1(~ demo_age + demo_weight+ demo_height+BMI + demo_gender + thc_chg + postConsumpTimeToTest + min_constriction_chg + AUC_chg + duration_chg + slope_chg + avg_end_percent_chg|user_cat, 
       data = pt.df)
```

```{r outliers, echo=FALSE}
## Remove known outliers
ps$outliers <- 0
# ps$outliers[ps$subject_id == "001-109" & ps$tp == "pre2"] <- 1 # Ben revised
ps$outliers[ps$subject_id == "001-060" & ps$tp == "post"] <- 1

ps <- ps[ps$outliers == 0, ]
```

Plot of PRE/POST for Right Eye
```{r rightEye_plots, echo=FALSE}
right.df <- ps[ps$eye == "Right", ]

# ggplot(right.df, aes(x=frame, y = pupil_size, group = subject_id, color = subject_id))+
#   geom_line(show.legend = FALSE)+
#   facet_grid(rows = vars(user_cat), cols = vars(tp))+
#   labs(title ="Pupil Size by Pre/Post and THC use category")+
#   theme_bw()

ggplot(right.df, aes(x=frame, y = percent_change, group = subject_id, color = subject_id))+
  geom_line(show.legend = FALSE)+
  facet_grid(rows = vars(user_cat), cols = vars(tp))+
  labs(title = "Percent Change by Pre/Post and THC use category")+
  theme_bw()
```

# Exploratory Analysis

```{r mean_sd_functions, include=FALSE}
right_0.df <- right.df[right.df$frame >= 0, c("subject_id", "tp", "user_cat", 
                                              "frame", "percent_change")]

right_0.df <- right_0.df[order(right_0.df$subject_id, right_0.df$tp, right_0.df$frame), ]

right_0.w <- reshape(right_0.df, 
                     timevar = "frame", 
                     idvar = c("subject_id", "tp", "user_cat"), 
                     direction = "wide")

rownames(right_0.w) <- paste0(right_0.w$subject_id, "_", right_0.w$tp)
pct_chg <- names(right_0.w[, 4:602])

mean_fxn <- as.data.frame(aggregate(right_0.w[, 4:602], 
                                    list(right_0.w$tp, 
                                         right_0.w$user_cat),
                                    FUN = function(x) mean(x, na.rm = T)))

mean_fxn.l <- reshape(mean_fxn, 
                      varying = pct_chg, 
                      v.names = "percent_change", 
                      timevar = "frame", 
                      times = pct_chg, 
                      direction = "long")
mean_fxn.l$frame <- as.numeric(gsub("percent_change.", "", mean_fxn.l$frame))
rownames(mean_fxn.l) <- NULL
mean_fxn.l$id <- NULL

names(mean_fxn.l)[names(mean_fxn.l) == "Group.1"] <- "tp"
names(mean_fxn.l)[names(mean_fxn.l) == "Group.2"] <- "user_cat"


mean_fxn.l$grp <- paste0(mean_fxn.l$tp, "_", mean_fxn.l$user_cat)

ggplot(mean_fxn.l, aes(x=frame, y = percent_change, group = grp,  
                       color = user_cat))+
  geom_line()+
  facet_grid(rows = vars(tp))+
  labs(title = "Average Percent Change by Pre/Post and THC use category")+
  theme_bw()

sd_fxn <- as.data.frame(aggregate(right_0.w[, 4:602], 
                                    list(right_0.w$tp, 
                                         right_0.w$user_cat),
                                    FUN = function(x) sd(x, na.rm = T)))

```

## Analytic Sample
Make sure datasets have the same participants and are truncated at frame 400

```{r analyticN, echo = F}
right_0.post <- right_0.df[right_0.df$tp == "post", ]
right_0.post <- right_0.post[right_0.post$frame <= 400, ]
post.ids <- unique(right_0.post$subject_id)

right_0.post <- right_0.post[order(right_0.post$subject_id, right_0.post$frame), ]
right_0.post$frame_char <- str_pad(right_0.post$frame, width = 3, side = c("left"), pad = "0")

right_0.post.w <- reshape(right_0.post[, c("subject_id","tp", "user_cat", "frame_char", "percent_change")], 
                          timevar = "frame_char", 
                          idvar = c("subject_id", "tp", "user_cat"), 
                          direction = "wide")

var.order <- c(names(right_0.post.w)[1:3], sort(names(right_0.post.w)[4:length(names(right_0.post.w))]))

right_0.post.w <- right_0.post.w[, var.order]

var.names <- names(right_0.post.w[4:ncol(right_0.post.w)])

right_0.post <- reshape(right_0.post.w, 
                        varying = var.names, 
                        v.names = "percent_change", 
                        timevar = "frame", 
                        times = 0:400,
                        idvar = c("subject_id", "tp", "user_cat"), 
                        direction = "long")


right_0.pt <- reshape(right_0.df[right_0.df$frame <= 400,], 
                      timevar = "tp", 
                      idvar = c("subject_id", "user_cat", "frame"), 
                      direction = "wide")

right_0.pt$wiPctChg <- right_0.pt$percent_change.post - right_0.pt$percent_change.pre2

right_0.pt <- right_0.pt[, c("subject_id", "user_cat", "frame", "wiPctChg")]

right_0.pt <- right_0.pt[order(right_0.pt$subject_id, right_0.pt$frame), ]
right_0.pt$frame_char <- str_pad(right_0.pt$frame, width = 3, side = c("left"), pad = "0")

right_0.pt.w <- reshape(right_0.pt[, c("subject_id","user_cat", "frame_char", "wiPctChg")], 
                     timevar = "frame_char", 
                     idvar = c("subject_id", "user_cat"), 
                     direction = "wide")
var.order2 <- c(names(right_0.pt.w)[1:2], sort(names(right_0.pt.w)[3:length(names(right_0.pt.w))]))

right_0.pt.w <- right_0.pt.w[, var.order2]

# remove rows where all data is missing (e.g. pt didn't have both pre and post)
allMissing <- rowSums(is.na(right_0.pt.w[, 3:403]))
rowsAllMissing <- names(allMissing)[allMissing == 401]

right_0.pt.w <- right_0.pt.w[!(rownames(right_0.pt.w) %in% rowsAllMissing), ]

rownames(right_0.pt.w) <- paste0(right_0.pt.w$subject_id, "_", right_0.pt.w$user_cat)

analytic.N <- post.ids[post.ids %in% right_0.pt.w$subject_id]

## Filter all datasets to analytic sample

right_0.df <- right_0.df[right_0.df$subject_id %in% analytic.N,]
right_0.post <- right_0.post[right_0.post$subject_id %in% analytic.N, ]
right_0.post.w <- right_0.post.w[right_0.post.w$subject_id %in% analytic.N ,]
row.names(right_0.post.w) <- right_0.post.w$subject_id

right_0.pt <- right_0.pt[right_0.pt$subject_id %in% analytic.N, ]
right_0.pt.w <- right_0.pt.w[right_0.pt.w$subject_id %in% analytic.N, ]
row.names(right_0.pt.w) <- right_0.pt.w$subject_id
```

No Consumption end time recorded for non-users
```{r tab1_analyticN, echo = F}
# Demographics Table by User Category
table1(~demo_age + demo_weight+ demo_height+BMI + demo_gender + thc_chg + postConsumpTimeToTest + min_constriction_chg + AUC_chg + duration_chg + slope_chg + avg_end_percent_chg|user_cat, 
       data = pt.df[pt.df$subject_id %in% analytic.N, ])
```


## Post Data and Within Person Difference (WPD) Trajectories 
```{r postPlot, echo = F}
post_traj <- ggplot(right_0.post, aes(x=frame, y = percent_change, group = subject_id, color = subject_id))+
             geom_line(show.legend = FALSE)+
             facet_grid(rows = vars(user_cat))+
             labs(title ="POST Data Percent Change by THC Use Category", 
                  y = "Percent Change")+
             theme_bw()

wi_traj <- ggplot(right_0.pt, aes(x=frame, y = wiPctChg, group = subject_id, color = subject_id))+
           geom_line(show.legend = FALSE)+
           facet_grid(rows = vars(user_cat))+
           labs(title ="WPD Percent Change by THC", 
                y = "WPD in Percent Change")+
           theme_bw()

grid.arrange(post_traj, wi_traj, nrow = 1)
```

*Non-user plot seems "centered" around 0 but still showing heterogeneity. Lots of heterogeneity across user-groups, but a mostly positive difference.*

```{r getLegend, echo = F}
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
```

# Analysis

## fPCA

### fPCA post data
Truncating to frame 400

```{r fPCA_post, echo = F}
right_0.fpca <- fpca.face(Y = as.matrix(right_0.post.w[, 4:404]), pve = 0.95, knots = 30, var = T)

right_0.fpca.pve <- round(right_0.fpca$evalues/sum(right_0.fpca$evalues)*100, 2)

mu <- right_0.fpca$mu
right_sd <- sqrt(right_0.fpca$evalues)
right_Phi <- right_0.fpca$efunctions

df_plot <- data.frame(frame = seq(0, 400, by = 1), 
                      mu = mu, 
                      errband1 = 2*right_Phi[, 1]*right_sd[1], 
                      errband2 = 2*right_Phi[, 2]*right_sd[2], 
                      errband3 = 2*right_Phi[, 3]*right_sd[3], 
                      errband4 = 2*right_Phi[, 4]*right_sd[4])

colors <- c("Pop.Mean" = "black", "+2SD" = "red", "-2SD" = "blue")

plot_pc1 <- ggplot(data = df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband1, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband1, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC1:", "% Var Explained:", right_0.fpca.pve[1]))+
  scale_color_manual(values = colors)+
  theme_bw()

plot_pc2 <- ggplot(data = df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband2, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband2, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC2:", "% Var Explained:", right_0.fpca.pve[2]))+
  scale_color_manual(values = colors)+
  theme_bw()

plot_pc3 <- ggplot(data = df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband3, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband3, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC3:", "% Var Explained:", right_0.fpca.pve[3]))+
  scale_color_manual(values = colors)+
  theme_bw()

plot_pc4 <- ggplot(data = df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband4, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband4, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC4:", "% Var Explained:", right_0.fpca.pve[4]))+
  scale_color_manual(values = colors)+
  theme_bw()

legend_postPC <- g_legend(plot_pc1)

grid.arrange(plot_pc1+theme(legend.position = "none"), 
             plot_pc2+theme(legend.position = "none"), 
             plot_pc3+theme(legend.position = "none"), 
             plot_pc4+theme(legend.position = "none"),
             legend_postPC,
             layout_matrix = rbind(c(1,2,5), c(3,4,5)),
             widths = c(10, 10, 3),
             nrow = 2, ncol = 3)
```

### fPCA Within Person Differences
```{r fPCA_withinPersonChange, echo=F}
right_0_wi.fpca <- fpca.face(Y = as.matrix(right_0.pt.w[, 3:403]), pve = 0.95, knots = 30, var = T)
# str(right_0.fpca)

# plot_shiny(right_0.fpca)
right_0_wi.fpca.pve <- round(right_0_wi.fpca$evalues/sum(right_0_wi.fpca$evalues)*100, 2)

mu <- right_0_wi.fpca$mu
right_sd <- sqrt(right_0_wi.fpca$evalues)
right_Phi <- right_0_wi.fpca$efunctions 

wi.df_plot <- data.frame(frame = seq(0, 400, by = 1), 
                      mu = mu, 
                      errband1 = 2*right_Phi[, 1]*right_sd[1], 
                      errband2 = 2*right_Phi[, 2]*right_sd[2], 
                      errband3 = 2*right_Phi[, 3]*right_sd[3], 
                      errband4 = 2*right_Phi[, 4]*right_sd[4], 
                      errband5 = 2*right_Phi[, 5]*right_sd[5], 
                      errband6 = 2*right_Phi[, 6]*right_sd[6])

colors <- c("Pop.Mean" = "black", "+2SD" = "red", "-2SD" = "blue")

wi.plot_pc1 <- ggplot(data = wi.df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband1, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband1, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC1:", "% Var Explained:", right_0_wi.fpca.pve[1]))+
  scale_color_manual(values = colors)+
  theme_bw()

wi.plot_pc2 <- ggplot(data = wi.df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband2, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband2, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC2:", "% Var Explained:", right_0_wi.fpca.pve[2]))+
  scale_color_manual(values = colors)+
  theme_bw()

wi.plot_pc3 <- ggplot(data = wi.df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband3, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband3, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC3:", "% Var Explained:", right_0_wi.fpca.pve[3]))+
  scale_color_manual(values = colors)+
  theme_bw()

wi.plot_pc4 <- ggplot(data = wi.df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband4, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband4, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC4:", "% Var Explained:", right_0_wi.fpca.pve[4]))+
  scale_color_manual(values = colors)+
  theme_bw()

wi.plot_pc5 <- ggplot(data = wi.df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband5, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband5, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC5:", "% Var Explained:", right_0_wi.fpca.pve[5]))+
  scale_color_manual(values = colors)+
  theme_bw()

wi.plot_pc6 <- ggplot(data = wi.df_plot, aes(x = frame, y = mu))+
  geom_line(aes(color = "Pop.Mean"))+
  geom_line(aes(x = frame, y = mu+errband6, color = "+2SD"))+
  geom_line(aes(x = frame, y = mu-errband6, color = "-2SD"))+
  labs(x = "frame", 
       y = "Percent Change", 
       color = "Legend", 
       title = paste("PC6:", "% Var Explained:", right_0_wi.fpca.pve[6]))+
  scale_color_manual(values = colors)+
  theme_bw()

legend_wiPC <- g_legend(wi.plot_pc1)

grid.arrange(wi.plot_pc1+theme(legend.position = "none"), 
             wi.plot_pc2+theme(legend.position = "none"), 
             wi.plot_pc3+theme(legend.position = "none"), 
             wi.plot_pc4+theme(legend.position = "none"),
             wi.plot_pc5+theme(legend.position = "none"),
             wi.plot_pc6+theme(legend.position = "none"),
             legend_wiPC,
             layout_matrix = rbind(c(1,2,7), c(3,4,7), c(5,6,7)),
             widths = c(10, 10, 3),
             nrow = 3, ncol = 3)
```


```{r wi_scores, echo = F}
wi.scores <- as.data.frame(right_0_wi.fpca$scores)
names(wi.scores) <- c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6")
wi.scores$subject_id <- rownames(wi.scores)

pt.wi.scores <- merge(wi.scores, pt.df, 
                      by = "subject_id", 
                      all.x = T)

pt.wi.scores$smoker <- ifelse(pt.wi.scores$user_cat %in% c("daily", "occasional"),1, 0)
```

## AUC: Predict Smoker vs Non-smoker

```{r fpca_post, echo = F}
post_right_0.fpca <- fpca.face(Y = as.matrix(right_0.post.w[, 4:404]), pve = 0.95, knots = 30, var = T)

post_right_0.fpca.pve <- round(post_right_0.fpca$evalues/sum(post_right_0.fpca$evalues)*100, 2)
# post_right_0.fpca.pve

post_right_scores <- as.data.frame(post_right_0.fpca$scores)
names(post_right_scores) <- c("PC1", "PC2", "PC3", "PC4")
post_right_scores$subject_id <- rownames(post_right_scores)

pt.scores <- merge(post_right_scores, pt.df, 
                   by = c("subject_id"), 
                   all.x = T)

pt.scores$smoker <- ifelse(pt.scores$user_cat %in% c("daily", "occasional"),1, 0)
```

```{r auc, echo = F}
m1 <- glm(smoker ~ PC1 + PC2 + PC3 +PC4, family = "binomial", data = pt.scores)
# summary(m1)
pt.scores$pred <- predict(m1, pt.scores)
pt.scores.roc <- roc(response = pt.scores$smoker, predictor = pt.scores$pred)
# pt.scores.roc
post.auc.pc4 <- pt.scores.roc$auc

plot.roc(pt.scores.roc, main = "Prediction of Smoker by PCs and Scalar Features in Post and WPD")

m1.scalar <- glm(smoker ~ min_constriction.post + AUC.post + slope.post, family = "binomial", data = pt.scores)
# summary(m1.scalar)
pt.scores$pred <- predict(m1.scalar, pt.scores)
pt.scalar.roc <- roc(response = pt.scores$smoker, predictor = pt.scores$pred)
# pt.scalar.roc
post.auc.scalar <- pt.scalar.roc$auc
plot.roc(pt.scalar.roc, main = "", add = T, col = "darkred")

auc.df <- data.frame("Label" = c("Post AUC PCs", "Post AUC Scalar Feat"), 
                     "AUC" = c(post.auc.pc4, post.auc.scalar), 
                     stringsAsFactors = F)

### WPD rocs
m5 <- glm(smoker ~ PC1 + PC2 + PC3 +PC4, family = "binomial", data = pt.wi.scores)
# summary(m5)

pt.wi.scores$pred_sameNPC <- predict(m5, pt.wi.scores)
pt.wi.scores.roc2 <- roc(response = pt.wi.scores$smoker, 
                     predictor = pt.wi.scores$pred_sameNPC)
# pt.wi.scores.roc2
wi.auc.pcs <- pt.wi.scores.roc2$auc
auc.df <- rbind(auc.df, c("W/I Diff AUC PCS", wi.auc.pcs))

plot.roc(pt.wi.scores.roc2, 
         col = "darkblue", 
         add = T)

m5.scalar <- glm(smoker ~ min_constriction_chg + AUC_chg + slope_chg, family = "binomial", data = pt.wi.scores)
# summary(m5.scalar)

pt.wi.scores$pred_wi_scalar <- predict(m5.scalar, pt.wi.scores)
pt.wi.scalar.roc2 <- roc(response = pt.wi.scores$smoker, 
                     predictor = pt.wi.scores$pred_wi_scalar)
# pt.wi.scalar.roc2
wi.auc.scalars <- pt.wi.scalar.roc2$auc
auc.df <- rbind(auc.df, c("W/I Diff AUC Scalar Feat", wi.auc.scalars))
plot.roc(pt.wi.scalar.roc2, 
         col = "darkgreen",
         add = T)
legend(0.5, 0.35, legend = c(paste0("post PCs AUC: ",round(post.auc.pc4, 3)),
                                 paste0("post Scalar Features AUC: ",round(post.auc.scalar, 3)), 
                                 paste0("WPD PCs AUC: ", round(wi.auc.pcs, 3)),
                                 paste0("WPD Scalar Features AUC: ", round(wi.auc.scalars, 3))),
       col = c("black", "darkred", "darkblue", "darkgreen"), 
       lty = rep(1, 4), cex = 0.8, box.lty=0)

```


```{r compare_auc, include = F}
auc.df$AUC <- round(as.numeric(auc.df$AUC), 3)

knitr::kable(auc.df)
```

## Association with Demog Variables

```{r demogAssoc_prep, echo = F}
pt.scores$male <- ifelse(pt.scores$demo_gender == "Male", 1, 0)
pt.scores$non_user <- ifelse(pt.scores$user_cat == "non-user", 1, 0)
pt.scores$daily <- ifelse(pt.scores$user_cat == "daily", 1, 0)
pt.scores$occasional <- ifelse(pt.scores$user_cat == "occasional", 1, 0)


pt.wi.scores$male <- ifelse(pt.wi.scores$demo_gender == "Male", 1, 0)
pt.wi.scores$non_user <- ifelse(pt.wi.scores$user_cat == "non-user", 1, 0)
pt.wi.scores$daily <- ifelse(pt.wi.scores$user_cat == "daily", 1, 0)
pt.wi.scores$occasional <- ifelse(pt.wi.scores$user_cat == "occasional", 1, 0)

```


```{r corrPC_scalar, echo = F, eval=F}
chart.Correlation(pt.scores[, c("PC1", "PC2", "PC3", "PC4", 
                                "min_constriction.post", "duration.post", 
                                "avg_end_percent.post", "slope.post", "AUC.post")])

```


```{r corrPC_demog, echo = F, eval=F}
chart.Correlation(pt.scores[, c("PC1", "PC2", "PC3", "PC4", 
                                "demo_age","male", "thc.post", "thc_chg", "BMI", 
                                "postConsumpTimeToTest", "smoker", 
                                "non_user", "daily", "occasional")])
```


```{r corrScalar_demog, echo = F, eval=F}
chart.Correlation(pt.scores[, c("min_constriction.post", "duration.post", 
                                "avg_end_percent.post", "slope.post", "AUC.post", 
                                "demo_age","male", "thc.post", "thc_chg", "BMI", 
                                "postConsumpTimeToTest", "smoker", 
                                "non_user", "daily", "occasional")])
```

Demographic Variables:

+ Age
+ Weight
+ Height
+ BMI 
+ Change in THC
+ Time from consumption to Post test
+ Sex

Results: 

Regression models regressed demographic variables on the 4 PCs for the post data. 

+ BMI negatively associated with PC2 adjusting for other PCs (p= 0.03)

Regression models regressed demographic variables on the 6 PCs for the WPD data.

+ Weight negatively associated with PC2 &  positively associated with PC5 adjusting for other PCs (p = 0.04 & 0.009)
+ Height positively associated with PC5 adjusting for other PCs (p = 0.006)
+ BMI positively associated with PC1 adjusting for other PCs (p= 0.049)
+ Male positively associated with PC5 adjusting for other PCs (p= 0.040)


```{r demogAssoc_post, echo = F}
post.age.m <- lm(demo_age ~ PC1 + PC2 + PC3 + PC4, data = pt.scores)
# summary(post.age.m)

post.wt.m <- lm(demo_weight ~ PC1 + PC2 + PC3 + PC4, data = pt.scores)
# summary(post.wt.m)

post.ht.m <- lm(demo_height ~ PC1 + PC2 + PC3 + PC4, data = pt.scores)
# summary(post.ht.m)

post.bmi.m <- lm(BMI ~ PC1 + PC2 + PC3 + PC4, data = pt.scores)
# summary(post.bmi.m)

post.thc.m <- lm(thc_chg ~ PC1 + PC2 + PC3 + PC4, data = pt.scores)
# summary(post.thc.m)

post.consump.m <- lm(postConsumpTimeToTest ~ PC1 + PC2 + PC3 + PC4, data = pt.scores)
# summary(post.consump.m)

post.male.m <- glm(male ~ PC1 + PC2 + PC3 + PC4, family = "binomial", data = pt.scores)
# summary(post.male.m)
```

```{r wi_corrPC_scalar, echo=F, eval = F}
chart.Correlation(pt.wi.scores[, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6",
                                "min_constriction_chg", "AUC_chg", "duration_chg",
                                "avg_end_percent_chg", "slope_chg")])
```

```{r wi_corrPC_demog, echo=F, eval = F}
chart.Correlation(pt.wi.scores[, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6",
                                "demo_age","male", "thc.post", "thc_chg", "BMI", 
                                "postConsumpTimeToTest", "smoker", 
                                "non_user", "daily", "occasional")])
```

```{r wi_corrScalar_demog, echo=F, eval = F }
chart.Correlation(pt.wi.scores[, c("min_constriction_chg", "AUC_chg", "duration_chg", 
                                   "avg_end_percent_chg", "slope_chg", 
                                "demo_age","male", "thc.post", "thc_chg", "BMI", 
                                "postConsumpTimeToTest", "smoker", 
                                "non_user", "daily", "occasional")])
```

```{r demogAssoc_wi, echo=F}
wi.age.m <- lm(demo_age ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, data = pt.wi.scores)
# summary(wi.age.m)

wi.wt.m <- lm(demo_weight ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, data = pt.wi.scores)
# summary(wi.wt.m)

wi.ht.m <- lm(demo_height ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, data = pt.wi.scores)
# summary(wi.ht.m)

wi.bmi.m <- lm(BMI ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, data = pt.wi.scores)
# summary(wi.bmi.m)

wi.thc.m <- lm(thc_chg ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, data = pt.wi.scores)
# summary(wi.thc.m)

wi.consump.m <- lm(postConsumpTimeToTest ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, data = pt.wi.scores)
# summary(wi.consump.m)

pt.wi.scores$male <- ifelse(pt.wi.scores$demo_gender == "Male", 1, 0)

wi.male.m <- glm(male ~ PC1 + PC2 + PC3 + PC4 +PC5 + PC6, family = "binomial", data = pt.wi.scores)
# summary(wi.male.m)
```

## Backward Stepwise Selection

```{r step_post, echo=F, results='hide'}
post_step.df <- pt.scores[, c("PC1", "PC2", "PC3", "PC4", "min_constriction.post", 
                              "duration.post", "avg_end_percent.post", 
                              "slope.post", "AUC.post", "smoker")]

m.post_step <- glm(smoker ~ PC1 + PC2 + PC3 + PC4, family = "binomial", data = post_step.df)
post_step.res <- step(m.post_step, direction = "backward")
# summary(post_step.res)
# post_step.res$coefficients[-1]

post_step.df$pred_post_step <- predict(post_step.res, pt.scores)

pt.scores.roc3 <- roc(response = post_step.df$smoker, 
                     predictor = post_step.df$pred_post_step)

# pt.scores.roc3
auc.specific <- data.frame("Post AUC PCs Step", pt.scores.roc3$auc)
auc.df <- rbind(auc.df, setNames(auc.specific, names(auc.df)))

plot.roc(pt.scores.roc3, 
         col = "black")

## Scalar
m.post_step.scalar <- glm(smoker ~ min_constriction.post + AUC.post  + 
                            avg_end_percent.post + slope.post, 
                          family = "binomial", data = post_step.df)
post_step.res <- step(m.post_step.scalar, direction = "backward")
# summary(post_step.res)
# post_step.res$coefficients[-1]
post_step.df$pred_post_step.scalar <- predict(post_step.res, pt.scores)

pt.scalar.roc3 <- roc(response = post_step.df$smoker, 
                     predictor = post_step.df$pred_post_step.scalar)

# pt.scalar.roc3
auc.specific <- data.frame("Post AUC Scalar Step", pt.scalar.roc3$auc)
auc.df <- rbind(auc.df, setNames(auc.specific, names(auc.df)))

plot.roc(pt.scalar.roc3, 
         col = "darkred", 
         add = T)

## Within Person differences
wi_step.df <- pt.wi.scores[, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", 
                               "min_constriction_chg", "AUC_chg", "duration_chg", 
                               "avg_end_percent_chg", "slope_chg", "smoker")]

m.wi_step <- glm(smoker ~ PC1 + PC2 + PC3 + PC4 +PC5 +PC6, family = "binomial", data = wi_step.df)
wi_step.res <- step(m.wi_step, direction = "backward")
# wi_step.res$coeffcients[-1]
# summary(wi_step.res)

wi_step.df$pred_wi_step <- predict(wi_step.res, pt.wi.scores)
pt.wi.scores.roc4 <- roc(response = wi_step.df$smoker, 
                     predictor = wi_step.df$pred_wi_step)

# pt.wi.scores.roc4
auc.specific <- data.frame("W/I Diff AUC PCs Step", pt.wi.scores.roc4$auc)
auc.df <- rbind(auc.df, setNames(auc.specific, names(auc.df)))

plot.roc(pt.wi.scores.roc4,
         col = "darkblue", 
         add = T)

### Scalar Features
m.wi_step.scalar <- glm(smoker ~ min_constriction_chg + AUC_chg 
                 + avg_end_percent_chg + slope_chg,
                 family = "binomial", data = wi_step.df)

wi_step.res <- step(m.wi_step.scalar, direction = "backward")
# summary(wi_step.res)
# wi_step.res$coefficients[-1]

wi_step.df$pred_wi_step.scalar <- predict(wi_step.res, pt.wi.scores)
pt.wi.scalar.roc4 <- roc(response = wi_step.df$smoker, 
                     predictor = wi_step.df$pred_wi_step.scalar)

# pt.wi.scalar.roc4
auc.specific <- data.frame("W/I Diff AUC Scalar Step", pt.wi.scalar.roc4$auc)
auc.df <- rbind(auc.df, setNames(auc.specific, names(auc.df)))

plot.roc(pt.wi.scalar.roc4, 
         col = "darkgreen", 
         add = T)
legend(0.5, 0.35, legend = c(paste0("post PCs AUC: ",round(pt.scores.roc3$auc, 3)),
                                 paste0("post Scalar Features AUC: ",round(pt.scalar.roc3$auc, 3)), 
                                 paste0("WPD PCs AUC: ", round(pt.wi.scores.roc4$auc, 3)),
                                 paste0("WPD Scalar Features AUC: ", round(pt.wi.scalar.roc4$auc, 3))),
       col = c("black", "darkred", "darkblue", "darkgreen"), 
       lty = rep(1, 4), cex = 0.8, box.lty=0)
```

+ Post PCs Selected: PC3 and PC4
+ Post Scalar Features Selected: Minimal Constriction and AUC
+ WPD PCs Selected: PC1, PC2, PC3, PC5
+ WPD Scalar Features Selected: Change in minimal constriction

## Function on Scalar Regression

```{r pt_analytic, echo = F}
pt.analytic.df <- pt.df[pt.df$subject_id %in% analytic.N, ]
pt.analytic.df$occasional <- ifelse(pt.analytic.df$user_cat == "occasional", 1, 0)
pt.analytic.df$daily  <- ifelse(pt.analytic.df$user_cat == "daily", 1, 0)


pt.analytic.df <- pt.analytic.df[order(pt.analytic.df$subject_id), ]
```


```{r post_gam, echo = F}
right_0.gam.df <- merge(right_0.post, pt.analytic.df, 
                        by = c("subject_id", "user_cat"))

right_0.gam.df <- right_0.gam.df[order(right_0.gam.df$subject_id, right_0.gam.df$frame), ]
right_0.gam.df$sind <- right_0.gam.df$frame/400


m.post_gam <- mgcv::gam(percent_change ~ s(sind, k=30, bs="cr") + s(sind, by=occasional, k=30, bs = "cr") + s(sind, by=daily, k=30, bs = "cr"), 
                  data = right_0.gam.df, method = "REML")

## Create a matrix of residuals
post_gam.resid <- cbind(right_0.gam.df[!(is.na(right_0.gam.df$percent_change)), c("subject_id", "frame")], m.post_gam$residuals)
names(post_gam.resid) <- c("subject_id", "frame", "resid")
post_gam.resid$frame_char <- str_pad(post_gam.resid$frame, width = 3, side = "left", pad = "0")

resid_mat <- reshape(post_gam.resid[, c("subject_id", "frame_char", "resid")], 
                     timevar = "frame_char", 
                     idvar = "subject_id", 
                     direction = "wide")
rownames(resid_mat) <- resid_mat$subject_id
resid_mat$subject_id <- NULL
resid_names <- names(resid_mat)
resid_names <- sort(resid_names)
resid_mat <- resid_mat[, resid_names]

resid_mat <- as.matrix(resid_mat)

post_gam.fpca <- fpca.face(resid_mat, knots = 15)

#Create data frame of eigen functions and attach to original data
Phi_mat <- as.data.frame(post_gam.fpca$efunctions)
colnames(Phi_mat) <- paste0("Phi", seq(1, post_gam.fpca$npc))
Phi_mat$frame <- as.numeric(rownames(Phi_mat)) -1

right_0.gam.df <- merge(right_0.gam.df, Phi_mat, 
                        by = "frame", 
                        all.x = T)
right_0.gam.df <- right_0.gam.df[order(right_0.gam.df$subject_id, right_0.gam.df$frame), ]
right_0.gam.df$subject_id <- as.factor(right_0.gam.df$subject_id)

post_gam.fri <- mgcv::bam(percent_change ~ 
                            s(sind, k=30, bs="cr") + 
                            s(sind, by=occasional, k=30, bs = "cr") + 
                            s(sind, by=daily, k=30, bs = "cr") +
                            s(subject_id, by = Phi1, bs="re") + s(subject_id, by = Phi2, bs="re")+
                            s(subject_id, by = Phi3, bs="re") + s(subject_id, by = Phi4, bs="re"), 
                          method = "fREML", data = right_0.gam.df, discrete = TRUE)


# summary(post_gam.fri)
```

```{r plot_occ, echo = F, fig.show = 'hide'}
post_gam.coef <- post_gam.fri$coefficients
post_gam.cov <- vcov.gam(post_gam.fri)

df_pred_non <- data.frame("sind" = unique(right_0.gam.df$sind), "occasional" = 0, "daily" = 0, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = right_0.gam.df$subject_id[1])

lpmat_non <- predict(post_gam.fri, newdata=df_pred_non, se.fit=TRUE, type = "lpmatrix")


df_pred_occ <- data.frame("sind" = unique(right_0.gam.df$sind), "occasional" = 1, "daily" = 0, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = right_0.gam.df$subject_id[1])

lpmat_occ <- predict(post_gam.fri, newdata=df_pred_occ, se.fit=TRUE, type = "lpmatrix")

df_pred_dly <- data.frame("sind" = unique(right_0.gam.df$sind), "occasional" = 0, "daily" = 1, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = right_0.gam.df$subject_id[1])

lpmat_dly <- predict(post_gam.fri, newdata=df_pred_dly, se.fit=TRUE, type = "lpmatrix")


pred_df <- data.frame(frame = rep(seq(0, 400), 3),
                      user = c(rep("non-user", 401), rep("occasional", 401),rep("daily", 401)), 
                      mean = c(lpmat_non %*% post_gam.coef, 
                               lpmat_occ %*% post_gam.coef, lpmat_dly %*% post_gam.coef), 
                      se = c(sqrt(diag(lpmat_non %*% post_gam.cov %*% t(lpmat_non))), 
                             sqrt(diag(lpmat_occ %*% post_gam.cov %*% t(lpmat_occ))),
                             sqrt(diag(lpmat_dly %*% post_gam.cov %*% t(lpmat_dly)))), 
                      stringsAsFactors = F)

post_userProfile <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                    geom_line()+
                    # geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
                    # geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
                    labs(y = "Percent Change")+
                    theme_bw()

legend_userProfile <- g_legend(post_userProfile)

post_userProfile.se <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                    geom_line()+
                    geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
                    geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
                    labs(y = "")+
                    theme_bw()
                    
post_userTraj <- grid.arrange(arrangeGrob(post_userProfile+theme(legend.position = "none"), 
                                          post_userProfile.se+theme(legend.position = "none"), nrow = 1), 
                              legend_userProfile, nrow = 1, 
                              widths = c(30, 6))
```


```{r coef_effect_plots, echo = F, fig.show = 'hide'}
pred_f1 <- predict(post_gam.fri, newdata=df_pred_occ, se.fit=TRUE, type = "terms")
pred_f2 <- predict(post_gam.fri, newdata=df_pred_dly, se.fit=TRUE, type = "terms")

pred_coef_df <- data.frame(frame = seq(0, 400),
                           f0_hat = lpmat_non %*% post_gam.coef, 
                           f0_se = sqrt(diag(lpmat_non %*% post_gam.cov %*% t(lpmat_non))), 
                           f1_hat = pred_f1$fit[, 2], 
                           f1_se = pred_f1$se.fit[, 2], 
                           f2_hat = pred_f2$fit[, 3], 
                           f2_se = pred_f2$se.fit[, 3])

post_non_plt <- ggplot(data = pred_coef_df, aes(x=frame, y = f0_hat))+
                geom_line()+
                geom_line(aes(x = frame, y = f0_hat + 2*f0_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = f0_hat - 2*f0_se), linetype = "longdash")+
                labs(title = "Average Percent Change of a Non-user (Post)", ylab = "Percent Change")+
                theme_bw()

post_occ_plt <- ggplot(data = pred_coef_df, aes(x=frame, y = f1_hat))+
                geom_line()+
                geom_line(aes(x = frame, y = f1_hat + 2*f1_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = f1_hat - 2*f1_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = 0), col = "darkred")+
                labs(title = "Difference in Percent Change: Occasional and Non-user (Post)", 
                     y= "")+
                theme_bw()+
                scale_x_continuous(expand = c(0, 0))+
                theme(rect = element_rect(fill = "transparent"))

post_dly_plt <- ggplot(data = pred_coef_df, aes(x=frame, y = f2_hat))+
                geom_line()+
                geom_line(aes(x = frame, y = f2_hat + 2*f2_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = f2_hat - 2*f2_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = 0), col = "darkred")+
                labs(title = "Difference in Percent Change: Daily and Non-user (Post)", 
                     y = "")+
                theme_bw()+
                scale_x_continuous(expand = c(0, 0))+
                theme(rect = element_rect(fill = "transparent"))

ylab <- textGrob(label = "Difference in Percent Change", rot = 90, 
                 gp = gpar(fontsize = 12))
posval <- textGrob(label = sapply(strwrap("Less Constriction in User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))
negval <- textGrob(label = sapply(strwrap("More Constriction in User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))

post_occ_coef <- grid.arrange(ylab, posval, negval, post_occ_plt, 
                              ncol = 2, nrow = 2, 
                              layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))

post_dly_coef <- grid.arrange(ylab, posval, negval, post_dly_plt, 
                              ncol = 2, nrow = 2, 
                              layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))
```


```{r diff_dly_occ, echo = F, fig.show= 'hide'}
f2_f1 <- (lpmat_dly - lpmat_occ) %*% post_gam.coef
f2_f1.se <- sqrt(diag((lpmat_dly - lpmat_occ) %*% post_gam.cov %*% t((lpmat_dly - lpmat_occ))))

f2_f1_diff_df <- data.frame(frame = seq(0, 400), 
                            f2f1_diff = f2_f1, 
                            f2f1_se = f2_f1.se)

# ggplot(data = f2_f1_diff_df, aes(x=frame, y = f2f1_diff))+
#   geom_line()+
#   geom_line(aes(x = frame, y = f2f1_diff + 2*f2f1_se), linetype = "longdash")+
#   geom_line(aes(x = frame, y = f2f1_diff - 2*f2f1_se), linetype = "longdash")+
#   labs(title = "Difference between Average Daily vs Occasional User", ylab = "f2_hat - f1_hat")+
#   theme_bw()

post_dlyocc_plt <- ggplot(data = f2_f1_diff_df, aes(x=frame, y = f2f1_diff))+
                   geom_line()+
                   geom_line(aes(x = frame, y = f2f1_diff + 2*f2f1_se), linetype = "longdash")+
                   geom_line(aes(x = frame, y = f2f1_diff - 2*f2f1_se), linetype = "longdash")+
                   geom_line(aes(x = frame, y = 0), col = "darkred")+
                   labs(title = "Difference in Percent Change: Daily vs Occasional User (Post)", 
                        y = "")+
                   theme_bw()+
                   scale_x_continuous(expand = c(0, 0))+
                   theme(rect = element_rect(fill = "transparent"))

ylab <- textGrob(label = "Difference in Percent Change", rot = 90, 
                 gp = gpar(fontsize = 12))
posval <- textGrob(label = sapply(strwrap("Less Constriction in Daily User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, just = "bottom", 
                 gp = gpar(fontsize = 8))
negval <- textGrob(label = sapply(strwrap("More Constriction in Daily User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))

post_dlyocc_coef <- grid.arrange(ylab, posval, negval, post_dlyocc_plt, 
                                 ncol = 2, nrow = 2, 
                                 layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))
```

```{r smokervnon , echo = F}
smoker.gam.df <- merge(right_0.post, pt.analytic.df, 
                        by = c("subject_id", "user_cat"))
smoker.gam.df$sind <- smoker.gam.df$frame/400
smoker.gam.df$smoker <- ifelse(smoker.gam.df$user_cat == "non-user", 0, 1)

m.post_smoker_gam <- mgcv::gam(percent_change ~ s(sind, k=30, bs="cr") + 
                                 s(sind, by=smoker, k=30, bs = "cr"), 
                  data = smoker.gam.df, method = "REML")

## Create a matrix of residuals
post_smoker_gam.resid <- cbind(smoker.gam.df[!(is.na(smoker.gam.df$percent_change)), c("subject_id", "frame")], m.post_smoker_gam$residuals)
names(post_smoker_gam.resid) <- c("subject_id", "frame", "resid")

post_smoker_gam.resid$frame_char <- str_pad(post_smoker_gam.resid$frame, width = 3, side = "left", pad = "0")

resid_mat <- reshape(post_smoker_gam.resid[, c("subject_id", "frame_char", "resid")], 
                     timevar = "frame_char", 
                     idvar = "subject_id", 
                     direction = "wide")

rownames(resid_mat) <- resid_mat$subject_id
resid_mat$subject_id <- NULL
resid_names <- names(resid_mat)
resid_names <- sort(resid_names)
resid_mat <- resid_mat[, resid_names]

resid_mat <- as.matrix(resid_mat)


post_smoker_gam.fpca <- fpca.face(resid_mat, knots = 15)

#Create data frame of eigen functions and attach to original data
Phi_mat <- as.data.frame(post_smoker_gam.fpca$efunctions)
colnames(Phi_mat) <- paste0("Phi", seq(1, post_smoker_gam.fpca$npc))
Phi_mat$frame <- as.numeric(rownames(Phi_mat)) -1

smoker.gam.df <- merge(smoker.gam.df, Phi_mat, 
                        by = "frame", 
                        all.x = T)
smoker.gam.df <- smoker.gam.df[order(smoker.gam.df$subject_id, smoker.gam.df$frame), ]
smoker.gam.df$subject_id <- as.factor(smoker.gam.df$subject_id)

post_smoker_gam.fri <- mgcv::bam(percent_change ~ 
                            s(sind, k=30, bs="cr") + 
                            s(sind, by=smoker, k=30, bs = "cr") +
                            s(subject_id, by = Phi1, bs="re") + s(subject_id, by = Phi2, bs="re")+
                            s(subject_id, by = Phi3, bs="re") + s(subject_id, by = Phi4, bs="re"), 
                          method = "fREML", data = smoker.gam.df, discrete = TRUE)


# summary(post_smoker_gam.fri)
```

```{r plot_smk, echo = F, fig.show = 'hide'}
post_smoker_gam.coef <- post_smoker_gam.fri$coefficients
post_smoker_gam.cov <- vcov.gam(post_smoker_gam.fri)

df_pred_non <- data.frame("sind" = unique(smoker.gam.df$sind), "smoker" =0, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = smoker.gam.df$subject_id[1])

lpmat_non <- predict(post_smoker_gam.fri, newdata=df_pred_non, se.fit=TRUE, type = "lpmatrix")


df_pred_smk <- data.frame("sind" = unique(smoker.gam.df$sind), "smoker" = 1, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = smoker.gam.df$subject_id[1])

lpmat_smk <- predict(post_smoker_gam.fri, newdata=df_pred_smk, se.fit=TRUE, type = "lpmatrix")

pred_df <- data.frame(frame = rep(seq(0, 400), 2),
                      user = c(rep("non-user", 401), rep("smoker", 401)), 
                      mean = c(lpmat_non %*% post_smoker_gam.coef, 
                               lpmat_smk %*% post_smoker_gam.coef), 
                      se = c(sqrt(diag(lpmat_non %*% post_smoker_gam.cov %*% t(lpmat_non))), 
                             sqrt(diag(lpmat_smk %*% post_smoker_gam.cov %*% t(lpmat_smk)))), 
                      stringsAsFactors = F)

post_smkProfile <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                   geom_line()+
                   scale_y_continuous(limits = c(min(pred_df$mean - 2*pred_df$se),
                                                 max(pred_df$mean+2*pred_df$se)))+
                   labs(y = "Percent Change")+
                   theme_bw()

legend_smkProfile <- g_legend(post_smkProfile)

post_smkProfile.se <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                      geom_line()+
                      geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
                      geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
                      scale_y_continuous(limits = c(min(pred_df$mean - 2*pred_df$se),
                                                    max(pred_df$mean+2*pred_df$se)))+
                      labs(y = "")+
                      theme_bw()

post_smk <- grid.arrange(arrangeGrob(post_smkProfile+theme(legend.position = "none"),
                                     post_smkProfile.se+theme(legend.position = "none"), nrow = 1), 
                         legend_smkProfile, nrow = 1, 
                         widths = c(30, 6))
```

```{r smk_coef_effect_plots, echo = F, fig.show = 'hide'}
# pred_f0 <- predict(post_smoker_gam.fri, newdata=df_pred_non, se.fit=TRUE, type = "iterms")
pred_f1 <- predict(post_smoker_gam.fri, newdata=df_pred_smk, se.fit=TRUE, type = "terms")

pred_coef_df <- data.frame(frame = seq(0, 400),
                           f0_hat = lpmat_non %*% post_smoker_gam.coef, 
                           f0_se = sqrt(diag(lpmat_non %*% post_smoker_gam.cov %*% t(lpmat_non))), 
                           f1_hat = pred_f1$fit[, 2], 
                           f1_se = pred_f1$se.fit[, 2])

# ggplot(data = pred_coef_df, aes(x=frame, y = f0_hat))+
#   geom_line()+
#   geom_line(aes(x = frame, y = f0_hat + 2*f0_se), linetype = "longdash")+
#   geom_line(aes(x = frame, y = f0_hat - 2*f0_se), linetype = "longdash")+
#   labs(title = "Average Response of a Non-user", ylab = "f0_hat")+
#   theme_bw()

post_smk_plt <- ggplot(data = pred_coef_df, aes(x=frame, y = f1_hat))+
                geom_line()+
                geom_line(aes(x = frame, y = f1_hat + 2*f1_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = f1_hat - 2*f1_se), linetype = "longdash")+
                geom_line(aes(x = frame, y = 0), col = "darkred")+
                labs(title = "Difference in Percent Change: Smoker and Non-smoker", 
                     y = "")+
                theme_bw()+
                scale_x_continuous(expand = c(0, 0))+
                theme(rect = element_rect(fill = "transparent"))

ylab <- textGrob(label = "Difference in Percent Change", rot = 90, 
                 gp = gpar(fontsize = 12))
posval <- textGrob(label = sapply(strwrap("Less Constriction in Smokers", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, just = "bottom", 
                 gp = gpar(fontsize = 8))
negval <- textGrob(label = sapply(strwrap("More Constriction in Smokers", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))

post_smk_coef <- grid.arrange(ylab, posval, negval, post_smk_plt, 
                              ncol = 2, nrow = 2, 
                              layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))
```

```{r wi_fosr_pffr, include = F, eval = F}
right_0.pt.w <- right_0.pt.w[order(right_0.pt.w$subject_id), ]

wi_pffr.df <- data.frame("subject_id" = pt.analytic.df$subject_id,
                          "occ_user" = pt.analytic.df$occasional, 
                          "daily_user" = pt.analytic.df$daily, 
                          "wi_pct_chg" = I(as.matrix(right_0.pt.w[, c(3:403)])))
wi_pffr.df$subject_id <- as.factor(wi_pffr.df$subject_id)

m.wi_pffr <- pffr(wi_pct_chg ~ occ_user + daily_user + s(subject_id, bs="re"),
                  data = wi_pffr.df,
                  algorithm = "bam", 
                  method = "fREML", 
                  discrete = T)

summary(m.wi_pffr)
par(mfrow = c(1,3))
plot(m.wi_pffr)
```


```{r wi_gam, echo = F}
right_0.pt.gam.df <- merge(right_0.pt, pt.analytic.df, 
                        by = c("subject_id", "user_cat"))
right_0.pt.gam.df$sind <- right_0.pt.gam.df$frame/400


m.wi_gam <- mgcv::gam(wiPctChg ~ s(sind, k=30, bs="cr") + s(sind, by=occasional, k=30, bs = "cr") + s(sind, by=daily, k=30, bs = "cr"), 
                  data = right_0.pt.gam.df, method = "REML")


## Create a matrix of residuals
wi_gam.resid <- cbind(right_0.pt.gam.df[!(is.na(right_0.pt.gam.df$wiPctChg)), c("subject_id", "frame")], m.wi_gam$residuals)
names(wi_gam.resid) <- c("subject_id", "frame", "resid")

wi_gam.resid$frame_char <- str_pad(wi_gam.resid$frame, width = 3, side = "left", pad = "0")


resid_mat <- reshape(wi_gam.resid[, c("subject_id", "frame_char", "resid")],
                     timevar = "frame_char",
                     idvar = "subject_id",
                     direction = "wide")
rownames(resid_mat) <- resid_mat$subject_id
resid_mat$subject_id <- NULL
resid_names <- names(resid_mat)
resid_names <- sort(resid_names)
resid_mat <- resid_mat[, resid_names]
resid_mat <- as.matrix(resid_mat)

wi_gam.fpca <- fpca.face(resid_mat, knots = 15)
# wi_gam.fpca$evalues/sum(wi_gam.fpca$evalues)

#Create data frame of eigen functions and attach to original data
Phi_mat <- as.data.frame(wi_gam.fpca$efunctions)
colnames(Phi_mat) <- paste0("Phi", seq(1, wi_gam.fpca$npc))
Phi_mat$frame <- as.numeric(rownames(Phi_mat)) -1

right_0.pt.gam.df <- merge(right_0.pt.gam.df, Phi_mat,
                        by = "frame",
                        all.x = T)
right_0.pt.gam.df <- right_0.pt.gam.df[order(right_0.pt.gam.df$subject_id, right_0.pt.gam.df$frame), ]
right_0.pt.gam.df$subject_id <- as.factor(right_0.pt.gam.df$subject_id)

wi_gam.fri <- mgcv::bam(wiPctChg ~ s(sind, k=30, bs="cr") + 
                          s(sind, by=occasional, k=30, bs = "cr") + 
                          s(sind, by=daily, k=30, bs = "cr")+
                          s(subject_id, by = Phi1, bs="re") + s(subject_id, by = Phi2, bs="re")+
                          s(subject_id, by = Phi3, bs="re") + s(subject_id, by = Phi4, bs="re")+
                          s(subject_id, by = Phi5, bs="re") + s(subject_id, by = Phi6, bs="re"),
                        method = "fREML", data = right_0.pt.gam.df, discrete = TRUE)
# summary(wi_gam.fri)
```

```{r wi.plots, echo = F, fig.show='hide'}
wi_gam.coef <- wi_gam.fri$coefficients
wi_gam.cov <- vcov.gam(wi_gam.fri)

df_pred_non <- data.frame("sind" = unique(right_0.pt.gam.df$sind), "occasional" = 0, "daily" = 0, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, "Phi5" = 0, "Phi6" = 0, 
                          "subject_id" = right_0.pt.gam.df$subject_id[1])

lpmat_non <- predict(wi_gam.fri, newdata=df_pred_non, se.fit=TRUE, type = "lpmatrix")


df_pred_occ <- data.frame("sind" = unique(right_0.pt.gam.df$sind), "occasional" = 1, "daily" = 0, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, "Phi5" = 0, "Phi6" = 0, 
                          "subject_id" = right_0.pt.gam.df$subject_id[1])

lpmat_occ <- predict(wi_gam.fri, newdata=df_pred_occ, se.fit=TRUE, type = "lpmatrix")

df_pred_dly <- data.frame("sind" = unique(right_0.pt.gam.df$sind), "occasional" = 0, "daily" = 1, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, "Phi5" = 0, "Phi6" = 0, 
                          "subject_id" = right_0.pt.gam.df$subject_id[1])

lpmat_dly <- predict(wi_gam.fri, newdata=df_pred_dly, se.fit=TRUE, type = "lpmatrix")


pred_df <- data.frame(frame = rep(seq(0, 400), 3),
                      user = c(rep("non-user", 401), rep("occasional", 401),rep("daily", 401)), 
                      mean = c(lpmat_non %*% wi_gam.coef, 
                               lpmat_occ %*% wi_gam.coef, lpmat_dly %*% wi_gam.coef), 
                      se = c(sqrt(diag(lpmat_non %*% wi_gam.cov %*% t(lpmat_non))), 
                             sqrt(diag(lpmat_occ %*% wi_gam.cov %*% t(lpmat_occ))),
                             sqrt(diag(lpmat_dly %*% wi_gam.cov %*% t(lpmat_dly)))), 
                      stringsAsFactors = F)

# ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
#   geom_line()+
#   geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
#   geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
#   theme_bw()

wi_userProfile <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                   geom_line()+
                   scale_y_continuous(limits = c(min(pred_df$mean - 2*pred_df$se),
                                                 max(pred_df$mean+2*pred_df$se)))+
                   labs(y = "Within Person Difference in Percent Change")+
                   theme_bw()

legend_userProfile <- g_legend(wi_userProfile)

wi_userProfile.se <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                      geom_line()+
                      geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
                      geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
                      scale_y_continuous(limits = c(min(pred_df$mean - 2*pred_df$se),
                                                    max(pred_df$mean+2*pred_df$se)))+
                      labs(y = "")+
                      theme_bw()

wi_userTraj <- grid.arrange(arrangeGrob(wi_userProfile+theme(legend.position = "none"), 
                         wi_userProfile.se+theme(legend.position = "none"), nrow = 1), 
                         legend_userProfile, nrow = 1, 
                         widths = c(30, 6))
```
### Plotting average trajectories of users
```{r wpd_traj, echo = F,  fig.height= 6, fig.width=20}
grid.arrange(post_userTraj, wi_userTraj, ncol = 2)
```


```{r wicoef_effect_plots, echo = F, fig.show = 'hide'}
pred_f1 <- predict(wi_gam.fri, newdata=df_pred_occ, se.fit=TRUE, type = "terms")
pred_f2 <- predict(wi_gam.fri, newdata=df_pred_dly, se.fit=TRUE, type = "terms")

pred_coef_df <- data.frame(frame = seq(0, 400),
                           f0_hat = lpmat_non %*% wi_gam.coef, 
                           f0_se = sqrt(diag(lpmat_non %*% wi_gam.cov %*% t(lpmat_non))), 
                           f1_hat = pred_f1$fit[, 2], 
                           f1_se = pred_f1$se.fit[, 2], 
                           f2_hat = pred_f2$fit[, 3], 
                           f2_se = pred_f2$se.fit[, 3])

# ggplot(data = pred_coef_df, aes(x=frame, y = f0_hat))+
#   geom_line()+
#   geom_line(aes(x = frame, y = f0_hat + 2*f0_se), linetype = "longdash")+
#   geom_line(aes(x = frame, y = f0_hat - 2*f0_se), linetype = "longdash")+
#   labs(title = "Average Response of a Non-user", ylab = "Within Person Difference in Percent Change")+
#   theme_bw()

wi_occ_plt <-  ggplot(data = pred_coef_df, aes(x=frame, y = f1_hat))+
               geom_line()+
               geom_line(aes(x = frame, y = f1_hat + 2*f1_se), linetype = "longdash")+
               geom_line(aes(x = frame, y = f1_hat - 2*f1_se), linetype = "longdash")+
               geom_line(aes(x = frame, y = 0), col = "darkred")+
               labs(title = "Difference in Within Person Difference (WPD): Occasional and Non-user", 
                    y = "")+
               theme_bw()+
               scale_x_continuous(expand = c(0, 0))+
               theme(rect = element_rect(fill = "transparent"))

wi_dly_plt <- ggplot(data = pred_coef_df, aes(x=frame, y = f2_hat))+
              geom_line()+
              geom_line(aes(x = frame, y = f2_hat + 2*f2_se), linetype = "longdash")+
              geom_line(aes(x = frame, y = f2_hat - 2*f2_se), linetype = "longdash")+
              geom_line(aes(x = frame, y = 0), col = "darkred")+
              labs(title = "Difference in Within Person Difference (WPD): Daily and Non-user", 
                   y = "")+
              theme_bw()+
              scale_x_continuous(expand = c(0, 0))+
              theme(rect = element_rect(fill = "transparent"))

ylab <- textGrob(label = "Difference in WPD in Percent Change", rot = 90, 
                 gp = gpar(fontsize = 12))
posval <- textGrob(label = sapply(strwrap("Less WPD Constriction in User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))
negval <- textGrob(label = sapply(strwrap("More WPD Constriction in User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))

wi_occ_coef <- grid.arrange(ylab, posval, negval, wi_occ_plt, 
                              ncol = 2, nrow = 2, 
                              layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))

wi_dly_coef <- grid.arrange(ylab, posval, negval, wi_dly_plt, 
                              ncol = 2, nrow = 2, 
                              layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))
```

### Difference between Occasional and Non-user

```{r occ_coef_plt, echo=FALSE, fig.height = 6, fig.width=15}
grid.arrange(post_occ_coef, wi_occ_coef, ncol = 2)
```

### Difference between Daily and Non-Users

```{r dly_coef_plt, echo=FALSE, fig.height = 6, fig.width=15}
grid.arrange(post_dly_coef, wi_dly_coef, ncol = 2)
```

```{r wi_diff_dly_occ, echo = F, fig.show= 'hide'}
f2_f1 <- (lpmat_dly - lpmat_occ) %*% wi_gam.coef
f2_f1.se <- sqrt(diag((lpmat_dly - lpmat_occ) %*% wi_gam.cov %*% t((lpmat_dly - lpmat_occ))))

f2_f1_diff_df <- data.frame(frame = seq(0, 400), 
                            f2f1_diff = f2_f1, 
                            f2f1_se = f2_f1.se)

wi_dlyocc_plt <- ggplot(data = f2_f1_diff_df, aes(x=frame, y = f2f1_diff))+
                   geom_line()+
                   geom_line(aes(x = frame, y = f2f1_diff + 2*f2f1_se), linetype = "longdash")+
                   geom_line(aes(x = frame, y = f2f1_diff - 2*f2f1_se), linetype = "longdash")+
                   geom_line(aes(x = frame, y = 0), col = "darkred")+
                   labs(title = "Difference in WPD in Percent Change: Daily vs Occasional User", 
                        y = "")+
                   theme_bw()+
                   scale_x_continuous(expand = c(0, 0))+
                   theme(rect = element_rect(fill = "transparent"))

ylab <- textGrob(label = "Difference in WPD in Percent Change", rot = 90, 
                 gp = gpar(fontsize = 12))
posval <- textGrob(label = sapply(strwrap("Less Constriction in Daily User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, just = "bottom", 
                 gp = gpar(fontsize = 8))
negval <- textGrob(label = sapply(strwrap("More Constriction in Daily User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))

wi_dlyocc_coef <- grid.arrange(ylab, posval, negval, post_dlyocc_plt, 
                                 ncol = 2, nrow = 2, 
                                 layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))
```

### Difference between Daily and Occasional Users

```{r dlyocc_coef_plt, echo = F, fig.height = 6, fig.width=15}
grid.arrange(post_dlyocc_coef, wi_dlyocc_coef, ncol = 2)
```

## Difference in Between Smoker vs Non-smokers

```{r wi_smokervnon, echo = F}
wi.smoker.gam.df <- merge(right_0.pt, pt.analytic.df, 
                        by = c("subject_id", "user_cat"))
wi.smoker.gam.df$sind <- wi.smoker.gam.df$frame/400
wi.smoker.gam.df$smoker <- ifelse(wi.smoker.gam.df$user_cat == "non-user", 0, 1)

m.wi_smoker_gam <- mgcv::gam(wiPctChg ~ s(sind, k=15, bs="cr") + 
                                 s(sind, by=smoker, k=15, bs = "cr"), 
                  data = wi.smoker.gam.df, method = "REML")

## Create a matrix of residuals
wi_smoker_gam.resid <- cbind(wi.smoker.gam.df[!(is.na(wi.smoker.gam.df$wiPctChg)), c("subject_id", "frame")], m.wi_smoker_gam$residuals)
names(wi_smoker_gam.resid) <- c("subject_id", "frame", "resid")

wi_smoker_gam.resid$frame_char <- str_pad(wi_smoker_gam.resid$frame, width = 3, side = "left", pad = "0")

resid_mat <- reshape(wi_smoker_gam.resid[, c("subject_id", "frame_char", "resid")], 
                     timevar = "frame_char", 
                     idvar = "subject_id", 
                     direction = "wide")

rownames(resid_mat) <- resid_mat$subject_id
resid_mat$subject_id <- NULL
resid_names <- names(resid_mat)
resid_names <- sort(resid_names)
resid_mat <- resid_mat[, resid_names]

resid_mat <- as.matrix(resid_mat)


wi_smoker_gam.fpca <- fpca.face(resid_mat, knots = 15)

#Create data frame of eigen functions and attach to original data
Phi_mat <- as.data.frame(wi_smoker_gam.fpca$efunctions)
colnames(Phi_mat) <- paste0("Phi", seq(1, wi_smoker_gam.fpca$npc))
Phi_mat$frame <- as.numeric(rownames(Phi_mat)) -1

wi.smoker.gam.df <- merge(wi.smoker.gam.df, Phi_mat, 
                        by = "frame", 
                        all.x = T)
wi.smoker.gam.df <- smoker.gam.df[order(wi.smoker.gam.df$subject_id, wi.smoker.gam.df$frame), ]
wi.smoker.gam.df$subject_id <- as.factor(wi.smoker.gam.df$subject_id)

wi_smoker_gam.fri <- mgcv::bam(percent_change ~ 
                            s(sind, k=15, bs="cr") + 
                            s(sind, by=smoker, k=15, bs = "cr") +
                            s(subject_id, by = Phi1, bs="re") + s(subject_id, by = Phi2, bs="re")+
                            s(subject_id, by = Phi3, bs="re") + s(subject_id, by = Phi4, bs="re"), 
                          method = "fREML", data = wi.smoker.gam.df, discrete = TRUE)


# summary(wi_smoker_gam.fri)
```


```{r wi_plot_smk, echo = F, fig.show = 'hide'}
wi_smoker_gam.coef <- wi_smoker_gam.fri$coefficients
wi_smoker_gam.cov <- vcov.gam(wi_smoker_gam.fri)

df_pred_non <- data.frame("sind" = sort(unique(wi.smoker.gam.df$sind)), "smoker" =0, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = wi.smoker.gam.df$subject_id[1])

lpmat_non <- predict(wi_smoker_gam.fri, newdata=df_pred_non, se.fit=TRUE, type = "lpmatrix")


df_pred_smk <- data.frame("sind" = sort(unique(wi.smoker.gam.df$sind)), "smoker" = 1, 
                            "Phi1" = 0, "Phi2" = 0, "Phi3" = 0, 
                            "Phi4" = 0, 
                          "subject_id" = wi.smoker.gam.df$subject_id[1])

lpmat_smk <- predict(wi_smoker_gam.fri, newdata=df_pred_smk, se.fit=TRUE, type = "lpmatrix")

pred_df <- data.frame(frame = rep(seq(0, 400), 2),
                      user = c(rep("non-user", 401), rep("smoker", 401)), 
                      mean = c(lpmat_non %*% wi_smoker_gam.coef, 
                               lpmat_smk %*% wi_smoker_gam.coef), 
                      se = c(sqrt(diag(lpmat_non %*% wi_smoker_gam.cov %*% t(lpmat_non))), 
                             sqrt(diag(lpmat_smk %*% wi_smoker_gam.cov %*% t(lpmat_smk)))), 
                      stringsAsFactors = F)

# ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
#   geom_line()+
#   geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
#   geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
#   theme_bw()

wi_smkProfile <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                   geom_line()+
                   scale_y_continuous(limits = c(min(pred_df$mean - 2*pred_df$se),
                                                 max(pred_df$mean+2*pred_df$se)))+
                   labs(y = "WPD in Percent Change")+
                   theme_bw()

legend_smkProfile <- g_legend(wi_smkProfile)

wi_smkProfile.se <- ggplot(data = pred_df, aes(x = frame, y = mean, group = user, col = user))+
                      geom_line()+
                      geom_line(aes(x=frame, y = mean + 2*se, group = user, col = user), linetype = "longdash")+
                      geom_line(aes(x=frame, y = mean - 2*se, group = user, col = user), linetype = "longdash")+
                      scale_y_continuous(limits = c(min(pred_df$mean - 2*pred_df$se),
                                                    max(pred_df$mean+2*pred_df$se)))+
                      labs(y = "")+
                      theme_bw()

wi_smk <- grid.arrange(arrangeGrob(wi_smkProfile+theme(legend.position = "none"), 
                                   wi_smkProfile.se+theme(legend.position = "none"), nrow = 1), 
                       legend_smkProfile, nrow = 1, 
                       widths = c(30, 6))
```

### Average Trajectory of Smoker vs Non-Smoker
```{r plt_smkTraj,echo = F, fig.height= 8, fig.width= 8}
grid.arrange(post_smk, wi_smk, nrow = 2)
```


```{r wi_smk_coef_effect_plots, echo= F, fig.show = 'hide'}
# pred_f0 <- predict(post_smoker_gam.fri, newdata=df_pred_non, se.fit=TRUE, type = "iterms")
pred_f1 <- predict(wi_smoker_gam.fri, newdata=df_pred_smk, se.fit=TRUE, type = "terms")

pred_coef_df <- data.frame(frame = seq(0, 400),
                           f0_hat = lpmat_non %*% wi_smoker_gam.coef, 
                           f0_se = sqrt(diag(lpmat_non %*% wi_smoker_gam.cov %*% t(lpmat_non))), 
                           f1_hat = pred_f1$fit[, 2], 
                           f1_se = pred_f1$se.fit[, 2])

# ggplot(data = pred_coef_df, aes(x=frame, y = f0_hat))+
#   geom_line()+
#   geom_line(aes(x = frame, y = f0_hat + 2*f0_se), linetype = "longdash")+
#   geom_line(aes(x = frame, y = f0_hat - 2*f0_se), linetype = "longdash")+
#   labs(title = "Average Within Person Difference in Percent Change of a Non-user", ylab = "f0_hat")+
#   theme_bw()

wi_smk_plt <- ggplot(data = pred_coef_df, aes(x=frame, y = f1_hat))+
              geom_line()+
              geom_line(aes(x = frame, y = f1_hat + 2*f1_se), linetype = "longdash")+
              geom_line(aes(x = frame, y = f1_hat - 2*f1_se), linetype = "longdash")+
              geom_line(aes(x = frame, y = 0), col = "darkred")+
              labs(title = "Difference in Within Person Difference (WPD) in Percent Change \nbetween Smoker and Non-smoker", 
                   y = "")+
              theme_bw()+
              scale_x_continuous(expand = c(0, 0))+
              theme(rect = element_rect(fill = "transparent"))

ylab <- textGrob(label = "Difference in WPD in Percent Change", rot = 90, 
                 gp = gpar(fontsize = 12))
posval <- textGrob(label = sapply(strwrap("Less WPD Constriction in User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))
negval <- textGrob(label = sapply(strwrap("More WPD Constriction in User", width = 15, simplify = F),
                                  paste, collapse = "\n"), rot = 0, 
                 gp = gpar(fontsize = 8))

wi_smk_coef <- grid.arrange(ylab, posval, negval, wi_smk_plt, 
                            ncol = 2, nrow = 2, 
                            layout_matrix = rbind(c(1, 2, rep(4, 15)), c(1, 3, rep(4, 15))))
```

### Smoker vs Non-smoker Effect
```{r smokerEffects, echo = F, fig.height = 4, fig.width=15}
grid.arrange(post_smk_coef , wi_smk_coef, nrow = 1)
```


# Appendix

## FoSR
$$ 
\begin{aligned}
Y_{ij}(t) &= f_0(t) + f_1(t)*I(user = occasional) + f_2(t)*I(user = daily) + b_i(t) + \epsilon_i(t)\\
& b_i(t) \sim GP(0, \Sigma_b) \\
& \epsilon_i(t) \sim N(0, \sigma_{\epsilon}^2) \\
&= \sum_{k=1}^K \xi_{0k}\phi_{0k}(t_i) + \sum_{k=1}^K \xi_{1k}\phi_{1k}(t_i)*I(user = occasional) + \sum_{k=1}^K \xi_{2k}\phi_{2k}(t_i)*I(user = daily) + \sum_{k=1}^K \xi_{ik}\phi_k(t_i) + \epsilon_i(t)\\
\end{aligned}
$$

### Mean function for non-user
$$
\begin{aligned}
Y_{ij}(t) &= f_0(t)\\
&=\sum_{k=1}^K \xi_{0k}\phi_{0k}(t_i) \\
&= \Phi_0\xi_0
\end{aligned}
$$

### Mean function for occasional user
$$
\begin{aligned}
Y_{ij}(t) &= f_0(t) + f_1(t) \\  
&=\sum_{k=1}^K \xi_{0k}\phi_{0k}(t_i) + \sum_{k=1}^K \xi_{1k}\phi_{1k}(t_i)\\
&= \Phi_0\xi_0 + \Phi_1\xi_1\\ 
&= [ \Phi_0, \Phi_1 ] \xi
\end{aligned}
$$

### Mean function for daily user
$$
\begin{aligned}
Y_{ij}(t) &= f_0(t) + f_2(t) \\  
&=\sum_{k=1}^K \xi_{0k}\phi_{0k}(t_i) + \sum_{k=1}^K \xi_{2k}\phi_{2k}(t_i) \\
&= \Phi_0\xi_0 + \Phi_2\xi_2\\
&= [ \Phi_0, \Phi_2 ] \xi
\end{aligned}
$$
